Method

### 2021-01-11 Rmd/01_method.Rmd
Code responses

library(tidyverse)
library(papaja)
library(gt)
library(knitr)

dat_participants <- read.csv("Data/participants.csv")

@@

We collected data from `r dat_raw$id_db %>% unique %>% length() %>% printnum()` bilinguals from the Metropolitan Area of Barcelona, between `r min(dat_raw$time_stamp, na.rm = TRUE) %>% format(., "%dth %B, %Y")` and `r max(dat_raw$time_stamp, na.rm = TRUE) %>% format(., "%dth %B, %Y")`. All families gave informed consent before participating. This study was approved by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from Hospital del Mar (Barcelona, Spain), code XXXXXXXXX.

We assessed toddlers' language profile asking parents for an estimated proportion of exposure to each language. We excluded participants with >10% exposure to a third language. Fig. \@ref(fig:lpdist) illustrates the distribution of participants across language profiles and ages.

```{r lpdist, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

dat_participants %>%
  filter(lp %in% "Bilingual") %>% 
  mutate(bilingualism = cut_interval(
    bilingualism, n = 6,
    labels = paste0(seq(0, 50, by = 10), "%"))) %>%
  ggplot(aes(x = age_bin, y = bilingualism)) +
  stat_bin_2d(colour = "white") +
  scale_x_discrete(guide = guide_axis(n.dodge = 1)) +
  scale_fill_viridis_c(option = "plasma") +
  labs(x = "Age (months)", y = "L2 exposure", fill = "Participants") +
  theme_classic() +
  theme(axis.text = element_text(colour = "black"))
```

@@

We implemented an on-line questionnaire using `formr` [@arslan2020], divided in three forms: (1) language questionnaire, (2) demographic survey, and (3) Catalan and Spanish vocabulary checklists. Vocabulary checklists consisted in two lists of words (one in Catalan and one in Spanish). Items in one language were translation equivalents of the items in the other (e.g., whenever *gos* [dog] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping. When there were two acceptable translation equivalents for a given word, we included both (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged them into the same one (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*). We included items from a diverse sample of `r distinct(dat_items, category) %>% nrow()` semantic/functional categories: Action words, adventures, animals, body parts, clothes, colour, descriptive words, food and drink, furniture and rooms, household items, on-line, outside, parts of animals, parts of things, people, pronouns, quantifiers, question words, time, toys, and vehicles. We discarded the following categories: adverbs, auxiliary words, connectives, interjections and games and routines. The Catalan inventory contained `r filter(dat_items, language=="Catalan") %>% pull(item) %>% unique() %>% length()` items (`r filter(dat_items, cognate_rater1, language=="Catalan") %>% pull(item) %>% unique() %>% length()` cognates, `r filter(dat_items, !cognate_rater1, language=="Catalan") %>% pull(item) %>% unique() %>% length()` non-cognates) and the Spanish inventory contained `r filter(dat_items, language=="Spanish") %>% pull(item) %>% unique() %>% length()` (`r filter(dat_items, cognate_rater1, language=="Spanish") %>% pull(item) %>% unique() %>% length()` cognates, `r filter(dat_items, !cognate_rater1, language=="Spanish") %>% pull(item) %>% unique() %>% length()` non-cognates).

For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand and say it, or did not understand or say it (marked by default). Participants filled a long or a short version of the questionnaire. Participants presented with the long version filled a list of 800 translation equivalents (800 items in Catalan and 800 items in Spanish), while participants presented with a short were randomly allocated into one of four list of items. Each list contained a different set of ~400 translation equivalents (~400 in Catalan, ~400 in Spanish). Semantic/functional categories with less than 16 items--thus resulting in less than four items after dividing it in four lists - were not divided in the short version of the questionnaire: all of their items were included in the four lists. Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions.

@@

```{r n_responses}
n_obs <- nrow(dat_responses)
sum_obs <- dat_responses %>%
  group_by(item) %>%
  summarise(n = n(), .groups = "drop") %>%
  summarise(
    mean = mean(n),
    sd = sd(n),
    min = min(n),
    max = max(n)
  )

@@

We gathered `r printnum(n_obs, digits = 0)` responses. Items received an average of `r printnum(sum_obs$mean, digits = 2)` (*SD* = `r printnum(sum_obs$sd, digits = 2)`, `r Min = printnum(sum_obs$max, digits = 2)`, Max = `r printnum(sum_obs$max, digits = 2)`. We modelled the probability of participants' being reported to understand/produce a translation equivalent adjusting by:

@@

dat_items %>%
  drop_na(cognate_rater1) %>%
  filter(include) %>%
  select(A, B, C, D, cognate_rater1, frequency, language) %>% 
  mutate(cognate_rater1 = ifelse(cognate_rater1, "Cognate", "Non-cognate")) %>%
  pivot_longer(c(A, B, C, D), names_to = c("list"), values_to = "n") %>% 
  filter(n) %>% 
  group_by(list, language, cognate_rater1) %>% 
  summarise(n = sum(n, na.rm = TRUE),
            mean = mean(frequency, na.rm = TRUE),
            sd = sd(frequency, na.rm = TRUE),
            min = min(frequency, na.rm = TRUE),
            max = max(frequency, na.rm = TRUE),
            .groups = "drop") %>%
  ungroup() %>% 
  rename(cognateness = cognate_rater1) %>%
  relocate(list, language, cognateness, n, mean, sd, min, max) %>%
  arrange(list, language, cognateness) %>%
  mutate(list = paste("List", list)) %>% 
  pivot_wider(names_from = cognateness, values_from = c(n, mean, sd, min, max)) %>%
  clean_names() %>%
  gt(groupname_col = c("language")) %>% 
  fmt_number(columns = contains("cognate"), decimals = 2) %>%
  fmt_number(columns = starts_with("n_"), decimals = 0) %>%
  tab_header(title = md("**Questionnaire and item properties**"),
             subtitle = "Participants were randomly allocated on of the the four lists, and completed both the Catalan and the Spanish versions.") %>% 
  tab_spanner(label = "Frequency (Cognates)", columns = vars("n_cognate", "mean_cognate", "sd_cognate", "min_cognate", "max_cognate")) %>% 
  tab_spanner(label = "Frequency (Non-cognates)", columns = contains("non_cognate")) %>% 
  cols_label(list = md("**List**"), 
             n_cognate = md("*N*"),
             mean_cognate = md("*Mean*"),
             sd_cognate = md("*SD*"),
             min_cognate = md("*Min*"),
             max_cognate = md("*Max*"),
             n_non_cognate = md("*N*"),
             mean_non_cognate = md("*Mean*"),
             sd_non_cognate = md("*SD*"),
             min_non_cognate = md("*Min*"),
             max_non_cognate = md("*Max*"))

>>>>

library(tidyverse)
library(papaja)
library(gt)
library(knitr)
library(here)

participants <- read.csv(here("Data", "participants.csv"))
items <- read.csv(here("Data", "items.csv"))
vocabulary <- read.csv(here("Data", "vocabulary.csv"))
responses <- read.csv(here("Data", "responses.csv"))

@@

```{r participants}
n_participants <- length(unique(participants$id))
dates <- range(participants$time_stamp, na.rm = TRUE)
sex <- participants %>% drop_na(sex) %>% count(sex) %>% pull(n) %>% set_names(c("female", "male"))

We collected data from `r printnum(n_participants)` bilinguals (`r sex[1]` female), from the Metropolitan Area of Barcelona, between `r format(dates[1], "%dth %B, %Y")` and `r format(dates[2], "%dth %B, %Y")`. All families gave informed consent before participating. This study was approved by the Comitè d'Ètica de la Investigació amb Medicaments (CEIm) from Hospital del Mar (Barcelona, Spain), code XXXXXXXXX.

We assessed toddlers' language profile asking parents for an estimated proportion of exposure to each language. We excluded participants with >10% exposure to a third language. Fig. \@ref(tab:lps) illustrates the distribution of participants across language profiles and ages.

```{r lps, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

participants %>%
  mutate(age = cut_width(age, width = 4, boundary = 0),
         doe = round(doe, 1)) %>%
  count(age, doe) %>% 
  mutate(doe = scales::percent(doe)) %>% 
  pivot_wider(names_from = doe, values_from = n) %>% 
  gt() %>% 
  tab_spanner(columns = 2:7, label = md("**Exposure to L2**")) %>% 
  fmt_missing(columns = everything(), missing_text = 0) %>% 
  cols_label(age = md("**Age (months)**"))

@@

```{r items}
pool <- pool %>% 
  drop_na(cognate)
n_categories <- distinct(pool, category) %>% nrow()
n_items <- length(unique(pool$item))
n_item_language <- count(pool, language) %>% pull(n) %>% set_names(c("catalan", "spanish"))
n_item_cognate <- count(pool, language, cognate) %>% 
  pull(n) %>% 
  set_names(c("catalan-cognate", "catalan-noncognate", "spanish-cognate", "spanish-noncognate"))

We implemented an on-line questionnaire using `formr` [@arslan2020], divided in three forms: a (1) language questionnaire, a (2) demographic survey, and a (3) Catalan and a Spanish vocabulary checklists. Vocabulary checklists followed a similar structure as the Oxford Communicative Developmental Inventory [@hamilton2000] and consisted in two lists of words, one in Catalan and one in Spanish). Items in one language were translation equivalents of the items in the other (e.g., whenever *gos* [dog] was included in the Catalan inventory, the word *perro* was included in the Spanish inventory), roughly following a one-to-one mapping. When there were two acceptable translation equivalents for a given word, we included both (e.g., Catalan *acabar* [*to finish*] and Spanish *acabar* and *terminar*), or merged them into the same one (e.g., Spanish *mono* [*monkey*] and Catalan *mono/mico*). We included items from a diverse sample of `r n_categories` semantic/functional categorie (see Appendix 1). We discarded the following categories: adverbs, auxiliary words, connectives, interjections and games and routines. The Catalan inventory contained `r n_item_language["catalan"]` items (`r n_item_cognate["catalan-cognate"]` cognates, `r n_item_cognate["catalan-noncognate"]` non-cognates) and the Spanish inventory contained `r n_item_language["spanish"]` (`r n_item_cognate["spanish-cognate"]` cognates, `r n_item_cognate["spanish-noncognate"]` non-cognates).

For each word in the vocabulary checklists, we asked parents to report whether their child was able to understand it, understand and say it, or did not understand or say it (marked by default). Participants filled a long or a short version of the questionnaire. Participants presented with the long version filled a list of 800 translation equivalents (800 items in Catalan and 800 items in Spanish), while participants presented with a short were randomly allocated into one of four list of items. Each list contained a different set of ~400 translation equivalents (~400 in Catalan, ~400 in Spanish). Semantic/functional categories with less than 16 items--thus resulting in less than four items after dividing it in four lists - were not divided in the short version of the questionnaire: all of their items were included in the four lists. Another subset of items that were part of the trial lists of some experiments in the lab were also included in all versions. Table 2 in Appendix 1 shows the distribution of items accross questionnaire versions.

@@

```{r n_responses}
n_obs <- nrow(responses)
sum_obs <- responses %>%
  group_by(te) %>%
  summarise(n = n(), .groups = "drop") %>%
  summarise(
    mean = mean(n),
    sd = sd(n),
    min = min(n),
    max = max(n)
  )

@@

We gathered `r printnum(n_obs, big.mark = ",")` responses. Translation equivalents received an average of `r printnum(sum_obs$mean, digits = 2)` (*SD* = `r printnum(sum_obs$sd, digits = 2)`, `r Min = printnum(sum_obs$min, digits = 2)`, Max = `r printnum(sum_obs$max, digits = 2)`), both languages summed together. We modelled the probability of participants' being reported to understand/produce a translation equivalent adjusting by:

@@

pool %>%
  filter(include) %>%
  select(version, cognate, frequency = frequency_zipf, language) %>% 
  mutate(cognate = ifelse(cognate, "Cognate", "Non-cognate")) %>%
  group_by(version, language, cognate) %>% 
  summarise(
    n = sum(n, na.rm = TRUE),
    mean = mean(frequency, na.rm = TRUE),
    sd = sd(frequency, na.rm = TRUE),
    min = min(frequency, na.rm = TRUE),
    max = max(frequency, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ungroup() %>% 
  relocate(version, language, cognate, n, mean, sd, min, max) %>%
  arrange(version, language, cognate) %>%
  mutate(version = collapse(version, ",")) %>% 
  pivot_wider(names_from = cognate, values_from = c(n, mean, sd, min, max)) %>%
  clean_names() %>%
  gt(groupname_col = c("language")) %>% 
  fmt_number(columns = contains("cognate"), decimals = 2) %>%
  fmt_number(columns = starts_with("n_"), decimals = 0) %>%
  tab_header(title = md("**Questionnaire and item properties**"),
             subtitle = "Participants were randomly allocated on of the the four lists, and completed both the Catalan and the Spanish versions.") %>% 
  tab_spanner(label = "Frequency (Cognates)", columns = vars("n_cognate", "mean_cognate", "sd_cognate", "min_cognate", "max_cognate")) %>% 
  tab_spanner(label = "Frequency (Non-cognates)", columns = contains("non_cognate")) %>% 
  cols_label(list = md("**List**"), 
             n_cognate = md("*N*"),
             mean_cognate = md("*Mean*"),
             sd_cognate = md("*SD*"),
             min_cognate = md("*Min*"),
             max_cognate = md("*Max*"),
             n_non_cognate = md("*N*"),
             mean_non_cognate = md("*Mean*"),
             sd_non_cognate = md("*SD*"),
             min_non_cognate = md("*Min*"),
             max_non_cognate = md("*Max*"))
